---
title: "3 CNSVS relating modules to external clinical traits"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MDD CNSVS
## 0 set up r session

```{r}
library(WGCNA)
options(stringsAsFactors=FALSE)
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/1 data input/CNSVS/MDDinfCNSVS1.RData")
lnames
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/2 module detection/02-MDD-networkConstruction.RData")
lnames
```

## 3a quantifying module trait associations

```{r}
# define numbers of genes and samples
nMarkers=ncol(MDDinfData5)
nSamples=nrow(MDDinfData5)

# recalculate MEs with color labels
MEs0=moduleEigengenes(MDDinfData5,moduleColors)$eigengenes
MEs=orderMEs(MEs0)
moduleTraitCor=cor(MEs,MDDcnsvs2,use="p")
moduleTraitPvalue=corPvalueStudent(moduleTraitCor,nSamples)

# display correlations and p values
textMatrix = paste(signif(moduleTraitCor,2), "\n(",
                   signif(moduleTraitPvalue,1),")",sep="")
dim(textMatrix)=dim(moduleTraitCor)
par(mar=c(6,8.5,3,3))

labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = c("memory", "psychomotor speed","reaction time", "complex attention", "cognitive flexibility", "processing speed", "executive function", "verbal memory", "visual memory", "social acuity", "reasoning domain", "working memory", "sustained attention", "simple attention", "motor speed"),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = greenWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```

## 3b gene relationship to trait and important modules: Gene Significance and Module Membership

Quantify associations of individual inflammatory markers with trait of interest by defining Gene Significance as the absolute value of the correlation between inflammatory marker and the trait. For each module we also define a quantitative measure of module membership MM as the correlation of the module eigengene and the gene expression profile. This allows us to quantify the similarity of all genes on the array to every module.

```{r}
# Define variable executive function containing the weight column of MDDcnsvs2
ef = as.data.frame(MDDcnsvs2$executive_functioning_domain_ss);
names(ef) = "executive function"

# names (colors) of the modules
modNames = substring(names(MEs), 3)

geneModuleMembership = as.data.frame(cor(MDDcnsvs2, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneTraitSignificance = as.data.frame(cor(MDDcnsvs2, ef, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(ef), sep="");
names(GSPvalue) = paste("p.GS.", names(ef), sep="");

```

## 3c intramodular analysis: identifying genes with high GS and MM

See if GS and MM are highly correlated, which means that markers highly significantly associated with a trait are the most important elements of modules associated with the trait.

```{r}
module = "turquoise"
column = match(module, modNames);
moduleGenes = moduleColors==module;

sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Marker significance for executive",
                   main = paste("Module membership vs. marker significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

```

## 3d summary output of network analysis results

```{r}
MDDmodules = data.frame(markers = names(MDDinfData5),
                        module = moduleColors)

write.csv(MDDmodules, file = "MDDmodules.csv")
```


# Control CNSVS
## 0 set up r session

```{r}
library(WGCNA)
options(stringsAsFactors=FALSE)
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/1 data input/CNSVS/CNTRLinfCNSVS1.RData")
lnames
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/2 module detection/02-Control-networkConstruction.RData")
lnames
```

## 3a quantifying module trait associations

```{r}
# define numbers of genes and samples
nMarkers=ncol(CntrlinfData5)
nSamples=nrow(CntrlinfData5)

# recalculate MEs with color labels
MEs0=moduleEigengenes(CntrlinfData5,moduleColors)$eigengenes
MEs=orderMEs(MEs0)
moduleTraitCor=cor(MEs,Cntrlcnsvs2,use="p")
moduleTraitPvalue=corPvalueStudent(moduleTraitCor,nSamples)

# display correlations and p values
textMatrix = paste(signif(moduleTraitCor,2), "\n(",
                   signif(moduleTraitPvalue,1),")",sep="")
dim(textMatrix)=dim(moduleTraitCor)
par(mar=c(6,8.5,3,3))

labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = c("memory", "psychomotor speed","reaction time", "complex attention", "cognitive flexibility", "processing speed", "executive function", "verbal memory", "visual memory", "social acuity", "reasoning domain", "working memory", "sustained attention", "simple attention", "motor speed"),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = greenWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```

## 3b gene relationship to trait and important modules: Gene Significance and Module Membership

Quantify associations of individual inflammatory markers with trait of interest by defining Gene Significance as the absolute value of the correlation between inflammatory marker and the trait. For each module we also define a quantitative measure of module membership MM as the correlation of the module eigengene and the gene expression profile. This allows us to quantify the similarity of all genes on the array to every module.

```{r}
# Define variable executive function containing the weight column of MDDcnsvs2
ef = as.data.frame(Cntrlcnsvs2$executive_functioning_domain_ss);
names(ef) = "executive function"

# names (colors) of the modules
modNames = substring(names(MEs), 3)

geneModuleMembership = as.data.frame(cor(Cntrlcnsvs2, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneTraitSignificance = as.data.frame(cor(Cntrlcnsvs2, ef, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(ef), sep="");
names(GSPvalue) = paste("p.GS.", names(ef), sep="");

```

## 3c intramodular analysis: identifying genes with high GS and MM

See if GS and MM are highly correlated, which means that markers highly significantly associated with a trait are the most important elements of modules associated with the trait.

```{r}
module = "blue"
column = match(module, modNames);
moduleGenes = moduleColors==module;

sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Marker significance for executive",
                   main = paste("Module membership vs. marker significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

```

## 3d summary output of network analysis results

```{r}
Cntrlmodules = data.frame(markers = names(CntrlinfData5),
                        module = moduleColors)

write.csv(Cntrlmodules, file = "Cntrlmodules.csv")
```
