---
title: "3 madrs"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MDD

## setup

```{r}
library(WGCNA)
options(stringsAsFactors=FALSE)
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Depression subcomponents/MDDinfMadrs.RData")
lnames
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/2 module detection/02-MDD-networkConstruction.RData")
lnames
```

## 3a quantifying module trait associations

```{r}
# define numbers of genes and samples
nMarkers=ncol(MDDinfData5)
nSamples=nrow(MDDinfData5)

# recalculate MEs with color labels
MEs0=moduleEigengenes(MDDinfData5,moduleColors)$eigengenes
MEs=orderMEs(MEs0)
moduleTraitCor=cor(MEs,MDDmadrs1,use="p")
moduleTraitPvalue=corPvalueStudent(moduleTraitCor,nSamples)

# display correlations and p values
textMatrix = paste(signif(moduleTraitCor,2), "\n(",
                   signif(moduleTraitPvalue,1),")",sep="")
dim(textMatrix)=dim(moduleTraitCor)
par(mar=c(6,8.5,3,3))

labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(MDDmadrs1),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```

## 3b gene relationship to trait and important modules: Gene Significance and Module Membership

```{r}
ef = as.data.frame(MDDmadrs1$`reported sadness`);
names(ef) = "reported sadness"

# names (colors) of the modules
modNames = substring(names(MEs), 3)

geneModuleMembership = as.data.frame(cor(MDDmadrs1, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneTraitSignificance = as.data.frame(cor(MDDmadrs1, ef, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(ef), sep="");
names(GSPvalue) = paste("p.GS.", names(ef), sep="");

```

## 3c intramodular analysis: identifying genes with high GS and MM

See if GS and MM are highly correlated, which means that markers highly significantly associated with a trait are the most important elements of modules associated with the trait.

```{r}
module = "blue"
column = match(module, modNames);
moduleGenes = moduleColors==module;

sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Marker significance for executive",
                   main = paste("Module membership vs. marker significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

```

## 3d visualize

visualize network of eigengenes

```{r}
# Recalculate module eigengenes
MEs = moduleEigengenes(MDDinfData5, moduleColors)$eigengenes
# Isolate weight from the clinical traits
ef = as.data.frame(MDDmadrs1$`reported sadness`);
names(ef) = "reported sadness"
# Add the weight to existing module eigengenes
MET = orderMEs(cbind(MEs, ef))
# Plot the relationships among the eigengenes and the trait
sizeGrWindow(5,7.5);
par(cex = 0.9)
plotEigengeneNetworks(MET, "", marDendro = c(0,4,1,2), marHeatmap = c(3,4,1,2), cex.lab = 0.8, xLabelsAngle
= 90)
```

# Control

## setup

```{r}
library(WGCNA)
options(stringsAsFactors=FALSE)
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Depression subcomponents/CntrlinfMadrs.RData")
lnames
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/2 module detection/02-Control-networkConstruction.RData")
lnames
```

## 3a quantifying module trait associations

```{r}
# define numbers of genes and samples
nMarkers=ncol(CntrlinfData5)
nSamples=nrow(CntrlinfData5)

# recalculate MEs with color labels
MEs0=moduleEigengenes(CntrlinfData5,moduleColors)$eigengenes
MEs=orderMEs(MEs0)
moduleTraitCor=cor(MEs,Cntrlmadrs1,use="p")
moduleTraitPvalue=corPvalueStudent(moduleTraitCor,nSamples)

# display correlations and p values
textMatrix = paste(signif(moduleTraitCor,2), "\n(",
                   signif(moduleTraitPvalue,1),")",sep="")
dim(textMatrix)=dim(moduleTraitCor)
par(mar=c(6,8.5,3,3))

labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(Cntrlmadrs1),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```

## 3b gene relationship to trait and important modules: Gene Significance and Module Membership

```{r}
ef = as.data.frame(Cntrlmadrs1$`reported sadness`);
names(ef) = "reported sadness"

# names (colors) of the modules
modNames = substring(names(MEs), 3)

geneModuleMembership = as.data.frame(cor(Cntrlmadrs1, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneTraitSignificance = as.data.frame(cor(Cntrlmadrs1, ef, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(ef), sep="");
names(GSPvalue) = paste("p.GS.", names(ef), sep="");

```

## 3c intramodular analysis: identifying genes with high GS and MM

See if GS and MM are highly correlated, which means that markers highly significantly associated with a trait are the most important elements of modules associated with the trait.

```{r}
module = "blue"
column = match(module, modNames);
moduleGenes = moduleColors==module;

sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Marker significance for executive",
                   main = paste("Module membership vs. marker significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

```

## 3d visualize

visualize network of eigengenes

```{r}
# Recalculate module eigengenes
MEs = moduleEigengenes(CntrlinfData5, moduleColors)$eigengenes
# Isolate weight from the clinical traits
ef = as.data.frame(Cntrlmadrs1$`reported sadness`);
names(ef) = "reported sadness"
# Add the weight to existing module eigengenes
MET = orderMEs(cbind(MEs, ef))
# Plot the relationships among the eigengenes and the trait
sizeGrWindow(5,7.5);
par(cex = 0.9)
plotEigengeneNetworks(MET, "", marDendro = c(0,4,1,2), marHeatmap = c(3,4,1,2), cex.lab = 0.8, xLabelsAngle
= 90)
```
