---
title: "1 inflammatory markers"
output:
  html_document: 
    message: FALSE
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install and load required packages
if (!requireNamespace("visdat", quietly = TRUE)) {
  install.packages("visdat")
}
if (!requireNamespace("mice", quietly = TRUE)) {
  install.packages("mice")
}

library(visdat)
library(mice)
```

# 1. Inflammatory markers

## 1.a MDD

The baseline MDD inflammatory marker data is contained in the file `CBN01_CYTORAWL_DATA_Z3_01_V01_TRTMT.csv` that comes with CAN-BIND-1.

```{r}
# load package
library(WGCNA)
# do not omit
options(stringsAsFactors = FALSE)
# read data set
MDDinfData = read.csv("/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Inflammatory markers/CBN01_CYTORAWL_DATA_Z3_01_V01_TRTMT.csv")
```

`MDDinfData` contains 211 subjects with 34 variables. Each row corresponds to a patient and column to a inflammatory marker. We now remove auxiliary data.

```{r}
MDDinfData1 = as.data.frame(MDDinfData[,-c(1:4,34)])
rownames(MDDinfData1) = MDDinfData$SUBJLABEL
```

We first check for inflammatory markers and subjects with too many missing values:

```{r}
gsg = goodSamplesGenes(MDDinfData1, verbose=3)
# gsg$goodGenes
# gsg$goodSamples
# false <- which(!gsg$goodSamples)
# print(false)
```

Visualize the missing data:

```{r}
vis_miss(MDDinfData1)
```

Remove patients that are missing all the inflammatory data.

```{r}
if (!gsg$allOK) {
  MDDinfData2 <- MDDinfData1[gsg$goodSamples, gsg$goodGenes]
}

# MDDinfData2 = MDDinfData1[!apply(MDDinfData1, 1, function(x) all(is.na(x))), ]
vis_miss(MDDinfData2)
```

Apply multiple imputation by chained equation MICE method to impute the rest of the missing data.

*Description:* creates multiple imputed datasets, performs analysis on each, and combines results

*Pros:* (1) accounts for uncertainty in imputations. (2) maintains variability and relationships in data. (3) handles MCAR and MAR missing data types.

*Cons:* (1) complex and computationally intensive. (2) requires careful consideration of model parameters

```{r, include = FALSE}
# perform MICE with 5 imputed datasets, predictive mean matching method, 50 max number of iterations, 500 seed for reproducibility
impute = mice(MDDinfData2, m = 5, method = 'pmm', maxit = 50, seed = 500)
#summary(impute)
MDDinfData3 = complete(impute)
vis_miss(MDDinfData3)
```

Now we have 205 patients with 29 inflammatory markers.

Cluster the patients to see if there are obvious outliers.

```{r}
# perform hierarchial clustering
sampleTree = hclust(dist(MDDinfData3), method = "average");

# plot sample tree
sizeGrWindow(12,9)
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

Choose a height cut that will remove the offending sample. 5000.

```{r}
# plot line to show cut
# abline(h=5000, col="red")
# determine cluster under the line
clust = cutreeStatic(sampleTree, cutHeight = 5000, minSize = 10)
# table(clust)
# clust 1 contains the samples we want to keep.
keepSamples = (clust==1)
MDDinfData3 = MDDinfData3[keepSamples, ]
nGenes = ncol(MDDinfData3)
nSamples = nrow(MDDinfData3)

# visualize
sampleTree_cleaned <- hclust(dist(MDDinfData3), method = "average")
plot(sampleTree_cleaned, main = "Sample Clustering After Outlier Removal", sub = "", xlab = "Samples", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

We will be removing 10 patients. We now have 195 patients with 29 inflammatory markers.

Z score the dataframe.

```{r}
MDDinfData4=as.data.frame(scale(MDDinfData3))
```


Remove outliers within each inflammatory marker

```{r, include = FALSE}
# boxplot
boxplot_data=list()
for(col in colnames(MDDinfData4)){
  boxplot_data[[col]]=boxplot(MDDinfData4[[col]],plot=FALSE)$out
}
boxplot(boxplot_data,las=2,main="Outliers in each marker", ylab="Values",col="lightblue")

# remove outliers
remove_outliers=function(x){
  mean_val=mean(x,na.rm=TRUE)
  sd_val=sd(x,na.rm=TRUE)
  lower_limit=mean_val-3*sd_val
  upper_limit=mean_val+3*sd_val
  x[which(x<lower_limit | x > upper_limit)]=NA
  return(x)
}
MDDinfData5=apply(MDDinfData4,2,remove_outliers)
MDDinfData5=as.data.frame(MDDinfData5)

# visualizing missing data
vis_miss(MDDinfData5)

# perform multiple imputation
# perform MICE with 5 imputed datasets, predictive mean matching method, 50 max number of iterations, 500 seed for reproducibility
impute = mice(MDDinfData5, m = 5, method = 'pmm', maxit = 50, seed = 500)
#summary(impute)
MDDinfData5 = complete(impute)
vis_miss(MDDinfData5)

```


### Ignore below
Transform data according to CAN-BIND-1 instructions.

```{r}
# log10 (il1b, il1ra, il4, il17, eotaxin, g-csf, mcp-1, tnfa, il13, ip10, fgf, rantes)
columns_to_transform <- c("Hu.IL.1b..39.", "Hu.IL.1ra..25.", "Hu.IL.4..52.", "Hu.IL.7..74.", "Hu.Eotaxin..43.", "Hu.G.CSF..57.", "Hu.MCP.1.MCAF...53.", "Hu.TNF.a..36.", "Hu.IL.13..51.", "Hu.IP.10..48.", "Hu.FGF.basic..44.", "Hu.RANTES..37.")
MDDinfData3[, columns_to_transform] <- log10(MDDinfData3[, columns_to_transform] + 1)

# x^0.25 (il6, il12, il7, vegf)
columns_to_transform <- c("Hu.IL.6..19.", "Hu.IL.12.p70...75.", "Hu.IL.7..74.", "Hu.VEGF..45.")
MDDinfData3[, columns_to_transform] <- MDDinfData3[, columns_to_transform]^0.25

# x^0.5 (il8, il9, il2)
columns_to_transform <- c("Hu.IL.8..54.", "Hu.IL.9..77.", "Hu.IL.2..38.")
MDDinfData3[, columns_to_transform] <- MDDinfData3[, columns_to_transform]^0.5

# log2 (il10, gm-csf, il5, il15, mip-1b)
columns_to_transform <- c("Hu.IL.10..56.", "Hu.GM.CSF..34.", "Hu.IL.5..33.", "Hu.IL.15..73.", "Hu.MIP.1b..18.")
MDDinfData3[, columns_to_transform] <- log2(MDDinfData3[, columns_to_transform]+1)

# x^0.33 (ifng, mip-1a)
columns_to_transform <- c("Hu.IFN.g..21.", "Hu.MCP.1.MCAF...53.")
MDDinfData3[, columns_to_transform] <- MDDinfData3[, columns_to_transform]^0.33

# x^0.08333 (PDGF)
columns_to_transform <- c("Hu.PDGF.bb..47.")
MDDinfData3[, columns_to_transform] <- MDDinfData3[, columns_to_transform]^0.08333
```

Visualize the distribution 

```{r}
# histogram
#num_cols <- 6  # Number of columns
#num_rows <- ceiling(ncol(MDDinfData4) / num_cols)  # Number of rows
#par(mfrow = c(num_rows, num_cols))

#for (col in colnames(MDDinfData4)) {
#  if (is.numeric(MDDinfData4[[col]])) {
#    # Create a histogram for the current column
#    hist(MDDinfData4[[col]], main = paste("Histogram of", col), xlab = col, ylab = #"Frequency", col = "lightblue", border = "black")
#  } else {
#    message(paste("Column", col, "is not numeric and will be skipped."))
#  }
#}

#par(mfrow = c(1, 1))

# qq-plot
#generate_qq_plot <- function(MDDinfData4, col_name) {
#  ggplot(MDDinfData4, aes(sample = MDDinfData4[[col_name]])) +
#    stat_qq() +
#    stat_qq_line() +
#    ggtitle(paste("QQ Plot of", col_name)) +
#    xlab("Theoretical Quantiles") +
#    ylab("Sample Quantiles") +
#    theme_minimal()
#}

#plot_list <- list()

#for (col in colnames(MDDinfData4)) {
#  if (is.numeric(MDDinfData4[[col]])) {
#    plot_list[[col]] <- generate_qq_plot(MDDinfData4, col)
#  } else {
#    message(paste("Column", col, "is not numeric and will be skipped."))
#  }
#}

#grid.arrange(grobs = plot_list, ncol = 6)

# shapiro wilk test
#num_columns <- ncol(MDDinfData4)
#significant_columns <- list()

#for (i in 1:num_columns) {
#  result <- shapiro.test(MDDinfData4[[i]])
#  if (result$p.value < 0.05) {
#    significant_columns[[i]] <- list(column_number = i, p_value = result$p.value)
#  }
#}

#cat("Columns with p < 0.05:\n")
#for (i in seq_along(significant_columns)) {
#  if (!is.null(significant_columns[[i]])) {
#    cat("Column Number:", significant_columns[[i]]$column_number, ", p-value:", #significant_columns[[i]]$p_value, "\n")
#  }
#}

#IFABP
#CRP
#Hu.IL.1b..39.
#Hu.IL.1ra..25.
#Hu.IL.2..38.
#Hu.IL.5..33.
#$Hu.IL.6..19.
#$Hu.IL.7..74.
#$Hu.IL.8..54.
#$Hu.IL.10..56.
#$Hu.IL.12.p70...75.
#$Hu.IL.13..51.
#$Hu.IL.15..73.
#$Hu.IL.17..76.
#$Hu.G.CSF..57.
#$Hu.GM.CSF..34.
#$Hu.IFN.g..21.
#$Hu.IP.10..48.
#$Hu.MCP.1.MCAF...53.
#$Hu.MIP.1a..55.
#$Hu.PDGF.bb..47.
#$Hu.MIP.1b..18.
#$Hu.RANTES..37.
#$Hu.VEGF..45.

```

Transform the data frames (ignore)

```{r}
#MDDinfData4=MDDinfData3
#MDDinfData4$IFABP=log10(MDDinfData3$IFABP)
#MDDinfData4$CRP=log10(MDDinfData3$CRP)
#MDDinfData4$Hu.IL.1b..39.=log10(MDDinfData3$Hu.IL.1b..39.)
#MDDinfData4$Hu.IL.1ra..25.=log10(MDDinfData3$Hu.IL.1ra..25.)
#MDDinfData4$Hu.IL.2..38.=log10(MDDinfData3$Hu.IL.2..38.)
#MDDinfData4$Hu.IL.5..33.=log(MDDinfData3$Hu.IL.5..33.)
#MDDinfData4$Hu.IL.6..19.=log10(MDDinfData3$Hu.IL.6..19.)
#MDDinfData4$Hu.IL.7..74.=log10(MDDinfData3$Hu.IL.7..74.)
#MDDinfData4$Hu.IL.8..54.=log10(MDDinfData3$Hu.IL.8..54.)
#MDDinfData4$Hu.IL.10..56.=log10(MDDinfData3$Hu.IL.10..56.)
#MDDinfData4$Hu.IL.12.p70...75.=log10(MDDinfData3$Hu.IL.12.p70...75.)
#MDDinfData4$Hu.IL.13..51.=log10(MDDinfData3$Hu.IL.13..51.)
#MDDinfData4$Hu.IL.15..73.=log10(MDDinfData3$Hu.IL.15..73.)
#MDDinfData4$Hu.IL.17..76.=log10(MDDinfData3$Hu.IL.17..76.)
#MDDinfData4$Hu.G.CSF..57.=log10(MDDinfData3$Hu.G.CSF..57.)
#MDDinfData4$Hu.GM.CSF..34.=log10(MDDinfData3$Hu.GM.CSF..34.)
#MDDinfData4$Hu.IFN.g..21.=log10(MDDinfData3$Hu.IFN.g..21.)
#MDDinfData4$Hu.IP.10..48.=log10(MDDinfData3$Hu.IP.10..48.)
#MDDinfData4$Hu.MCP.1.MCAF...53.=log10(MDDinfData3$Hu.MCP.1.MCAF...53.)
#MDDinfData4$Hu.MIP.1a..55.=log10(MDDinfData3$Hu.MIP.1a..55.)
#MDDinfData4$Hu.PDGF.bb..47.=log10(MDDinfData3$Hu.PDGF.bb..47.)
#MDDinfData4$Hu.MIP.1b..18.=log10(MDDinfData3$Hu.MIP.1b..18.)
#MDDinfData4$Hu.RANTES..37.=log10(MDDinfData3$Hu.RANTES..37.)
#MDDinfData4$Hu.VEGF..45.=log10(MDDinfData3$Hu.VEGF..45.)

```

Perform quantile normalization to remove variation

```{r}
# Quantile normalization function
#quantile_normalize <- function(MDDinfData3) {
#  MDDinfData4 <- normalize.quantiles(as.matrix(MDDinfData3))
#  MDDinfData4 <- as.data.frame(MDDinfData4)
#  rownames(MDDinfData4) <- rownames(MDDinfData3)
#  colnames(MDDinfData4) <- colnames(MDDinfData3)
#  return(MDDinfData4)
#}

# Applying quantile normalization
#MDDinfData4 <- quantile_normalize(MDDinfData3)

# qqplot
#generate_qq_plot <- function(MDDinfData4, col_name) {
#  ggplot(MDDinfData4, aes(sample = MDDinfData4[[col_name]])) +
#    stat_qq() +
#    stat_qq_line() +
#    ggtitle(paste("QQ Plot of", col_name)) +
#    xlab("Theoretical Quantiles") +
#    ylab("Sample Quantiles") +
#    theme_minimal()
#}

#plot_list <- list()

#for (col in colnames(MDDinfData4)) {
#  if (is.numeric(MDDinfData4[[col]])) {
#    plot_list[[col]] <- generate_qq_plot(MDDinfData4, col)
#  } else {
#    message(paste("Column", col, "is not numeric and will be skipped."))
#  }
#}

#grid.arrange(grobs = plot_list, ncol = 6)

```

## 1.b Control

The baseline control inflammatory marker data is contained in the file `CBN01_CYTORAWL_DATA_Z3_01_V01_CNTRL.csv` that comes with CAN-BIND-1.

```{r}
# load package
library(WGCNA)
# do not omit
options(stringsAsFactors = FALSE)
# read data set
CntrlinfData = read.csv("/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Inflammatory markers/CBN01_CYTORAWL_DATA_Z3_01_V01_CNTRL.csv")
```

`CntrlinfData` contains 112 subjects with 34 variables. Each row corresponds to a participant and column to a inflammatory marker. We now remove auxiliary data.

```{r}
CntrlinfData1 = as.data.frame(CntrlinfData[,-c(1:4,34)])
rownames(CntrlinfData1) = CntrlinfData$SUBJLABEL
```

We first check for inflammatory markers and subjects with too many missing values:

```{r}
gsg = goodSamplesGenes(CntrlinfData1, verbose=3)
# gsg$goodGenes
# gsg$goodSamples
# false <- which(!gsg$goodSamples)
# print(false)
```

Visualize the missing data:

```{r}
vis_miss(CntrlinfData1)
```

Remove patients that are missing all the inflammatory data.

```{r}
CntrlinfData2 = CntrlinfData1[!apply(CntrlinfData1, 1, function(x) all(is.na(x))), ]
#vis_miss(CntrlinfData2)
```

Apply multiple imputation by chained equation MICE method to impute the rest of the missing data.

*Description:* creates multiple imputed datasets, performs analysis on each, and combines results

*Pros:* (1) accounts for uncertainty in imputations. (2) maintains variability and relationships in data. (3) handles MCAR and MAR missing data types.

*Cons:* (1) complex and computationally intensive. (2) requires careful consideration of model parameters

```{r, include=FALSE}
# perform MICE with 5 imputed datasets, predictive mean matching method, 50 max number of iterations, 500 seed for reproducibility
impute = mice(CntrlinfData2, m = 5, method = 'pmm', maxit = 50, seed = 500)
#summary(impute)
CntrlinfData3 = complete(impute)
#vis_miss(CntrlinfData3)
```

Now we have 102 patients with 29 inflammatory markers.

Cluster the participants to see if there are obvious outliers.

```{r}
# perform hierarchial clustering
sampleTree = hclust(dist(CntrlinfData3), method = "average");

# plot sample tree
sizeGrWindow(12,9)
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

Choose a height cut that will remove the offending sample. 5000.

```{r}
# plot line to show cut
# abline(h=5000, col="red")
# determine cluster under the line
clust = cutreeStatic(sampleTree, cutHeight = 5000, minSize = 10)
# table(clust)
# clust 1 contains the samples we want to keep.
keepSamples = (clust==1)
CntrlinfData4 = CntrlinfData3[keepSamples, ]
nGenes = ncol(CntrlinfData4)
nSamples = nrow(CntrlinfData4)

# visualize
sampleTree_cleaned <- hclust(dist(CntrlinfData4), method = "average")
plot(sampleTree_cleaned, main = "Sample Clustering After Outlier Removal", sub = "", xlab = "Samples", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

We will be removing 8 patients. We now have 94 patients with 29 inflammatory markers.

Z score the dataframe.

```{r}
CntrlinfData4=as.data.frame(scale(CntrlinfData4))
```


Remove outliers within each inflammatory marker

```{r, include=FALSE}
# boxplot
boxplot_data=list()
for(col in colnames(CntrlinfData4)){
  boxplot_data[[col]]=boxplot(CntrlinfData4[[col]],plot=FALSE)$out
}
boxplot(boxplot_data,las=2,main="Outliers in each marker", ylab="Values",col="lightblue")

# remove outliers
remove_outliers=function(x){
  mean_val=mean(x,na.rm=TRUE)
  sd_val=sd(x,na.rm=TRUE)
  lower_limit=mean_val-3*sd_val
  upper_limit=mean_val+3*sd_val
  x[which(x<lower_limit | x > upper_limit)]=NA
  return(x)
}
CntrlinfData5=apply(CntrlinfData4,2,remove_outliers)
CntrlinfData5=as.data.frame(CntrlinfData5)

# visualizing missing data
vis_miss(CntrlinfData5)

# perform multiple imputation
# perform MICE with 5 imputed datasets, predictive mean matching method, 50 max number of iterations, 500 seed for reproducibility
impute = mice(CntrlinfData5, m = 5, method = 'pmm', maxit = 50, seed = 500)
#summary(impute)
CntrlinfData5 = complete(impute)
vis_miss(CntrlinfData5)

```


