---
title: "2 sex"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MDD male
## 0 setting up r session

```{r}
# load libraries
library(WGCNA)
options(stringsAsFactors = FALSE)
enableWGCNAThreads()

# load data saved in first step
lnames = load(file = "/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Sex/sex.RData")
lnames 
```


## 2a choosing the soft thresholding power: analysis of network topology

```{r}
# choose a set of soft thresholding powers
powers=c(c(1:10), seq(from=12, to=40, by=2))

# call network topology analysis function
sft = pickSoftThreshold(MDDinfM, powerVector = powers, verbose = 5)

# plot the result
sizeGrWindow(9,5)
par(mfrow=c(1,2))
cex1=0.9

# scale free topology fit index as a function of the soft thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```

## 2b one step network construction and module detection

```{r}
# automatic construction is not working -- see below for step by step construction
net = blockwiseModules(MDDinfM, power = 6,
                       TOMType = "unsigned", minModuleSize = 1,
                       reassignThreshold = 0, mergeCutHeight = 0.25,
                       numericLabels = TRUE, pamRespectsDendro = FALSE,
                       saveTOMs = TRUE,
                       saveTOMFileBase = "TOM", 
                       verbose = 3)

# module sizes and number of modules
table(net$colors)
```

create a table of soft thresholding powers

```{r}
# Initialize an empty list to store the lengths of net$MEs for each power
MEs_lengths <- numeric(20)

# Loop through powers from 1 to 20
for (power in 1:20) {
  # Run blockwiseModules for the current power
  net <- blockwiseModules(MDDinfM, power = power,
                           TOMType = "unsigned", minModuleSize = 1,
                           reassignThreshold = 0, mergeCutHeight = 0.25,
                           numericLabels = TRUE, pamRespectsDendro = FALSE,
                           saveTOMs = TRUE,
                           saveTOMFileBase = paste0("TOM_", power), 
                           verbose = 3)
  
  # Store the length of net$MEs for the current power
  MEs_lengths[power] <- length(net$MEs)
}

# Create a data frame to store the lengths of net$MEs for each power
MEs_lengths_table <- data.frame(Power = 1:20, Length_of_MEs = MEs_lengths)

# Display the table
print(MEs_lengths_table)
```

We plot the hierarchical clustering dendrogram for module identification.


```{r}
# open a graphics window
sizeGrWindow(12, 9)
# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                    "Module colors",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
```

## 2c step by step construction
```{r}
# calculate adjacencies
softPower=10
adjacency=adjacency(MDDinfM,power=softPower)

# topological overlap matrix
TOM=TOMsimilarity(adjacency)
dissTOM=1-TOM

# clustering using TOM
geneTree=hclust(as.dist(dissTOM),method="average")
sizeGrWindow(12,9)
plot(geneTree,xlab="",sub="",main="Marker clustering on TOM based dissimilarity", labels=FALSE, hang=0.04)

# module identification using dynamic tree cut
minModuleSize=1
dynamicMods=cutreeDynamic(dendro=geneTree, distM=dissTOM, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize)
table(dynamicMods)

# convert numeric labels to colors
dynamicColors=labels2colors(dynamicMods)
table(dynamicColors)

# plot dendrograms and colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree,dynamicColors,"Dynamic Tree Cut", dendroLabels = FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05, main="Gene dendrogram and module colors")

# merging modules whose expression profiles are similar
MEList = moduleEigengenes(MDDinfM,colors=dynamicColors)
MEs = MEList$eigengenes
MEDiss = 1-cor(MEs)
METree = hclust(as.dist(MEDiss), method="average")
sizeGrWindow(7,6)
plot(METree, main="Clustering of module eigengenes", xlab="", sub="")

# choose height cut
MEDISSThres=0.25
merge=mergeCloseModules(MDDinfM,dynamicColors,cutHeight=MEDISSThres,verbose=3)
mergedColors=merge$colors
mergedMEs = merge$newMEs

sizeGrWindow(12,9)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),c("Dynamic Tree Cut", "Merged dynamic"), dendroLabels=FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05)

# save
moduleColors=mergedColors
colorOrder=c("grey",standardColors(50))
moduleLabels=match(moduleColors,colorOrder)-1
MEs=merge$newMEs
save(MEs, moduleLabels,moduleColors,geneTree,file="02-MDDinfM-networkConstruction.RData")
```

## Module output
```{r}
MDDMmodules = data.frame(markers = names(MDDinfM),
                        module = moduleColors)

write.csv(MDDMmodules, file = "MDDM.csv")
```




# MDD female
## 0 setting up r session

```{r}
# load libraries
library(WGCNA)
options(stringsAsFactors = FALSE)
enableWGCNAThreads()

# load data saved in first step
lnames = load(file = "/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Sex/sex.RData")
lnames 
```


## 2a choosing the soft thresholding power: analysis of network topology

```{r}
# choose a set of soft thresholding powers
powers=c(c(1:10), seq(from=12, to=40, by=2))

# call network topology analysis function
sft = pickSoftThreshold(MDDinfF, powerVector = powers, verbose = 5)

# plot the result
sizeGrWindow(9,5)
par(mfrow=c(1,2))
cex1=0.9

# scale free topology fit index as a function of the soft thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```

## 2b one step network construction and module detection

```{r}
# automatic construction is not working -- see below for step by step construction
net = blockwiseModules(MDDinfF, power = 10,
                       TOMType = "unsigned", minModuleSize = 1,
                       reassignThreshold = 0, mergeCutHeight = 0.25,
                       numericLabels = TRUE, pamRespectsDendro = FALSE,
                       saveTOMs = TRUE,
                       saveTOMFileBase = "TOM", 
                       verbose = 3)

# module sizes and number of modules
table(net$colors)
```

create a table of soft thresholding powers

```{r}
# Initialize an empty list to store the lengths of net$MEs for each power
MEs_lengths <- numeric(20)

# Loop through powers from 1 to 20
for (power in 1:20) {
  # Run blockwiseModules for the current power
  net <- blockwiseModules(MDDinfF, power = power,
                           TOMType = "unsigned", minModuleSize = 1,
                           reassignThreshold = 0, mergeCutHeight = 0.25,
                           numericLabels = TRUE, pamRespectsDendro = FALSE,
                           saveTOMs = TRUE,
                           saveTOMFileBase = paste0("TOM_", power), 
                           verbose = 3)
  
  # Store the length of net$MEs for the current power
  MEs_lengths[power] <- length(net$MEs)
}

# Create a data frame to store the lengths of net$MEs for each power
MEs_lengths_table <- data.frame(Power = 1:20, Length_of_MEs = MEs_lengths)

# Display the table
print(MEs_lengths_table)
```

We plot the hierarchical clustering dendrogram for module identification.


```{r}
# open a graphics window
sizeGrWindow(12, 9)
# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                    "Module colors",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
```

## 2c step by step construction
```{r}
# calculate adjacencies
softPower=10
adjacency=adjacency(MDDinfF,power=softPower)

# topological overlap matrix
TOM=TOMsimilarity(adjacency)
dissTOM=1-TOM

# clustering using TOM
geneTree=hclust(as.dist(dissTOM),method="average")
sizeGrWindow(12,9)
plot(geneTree,xlab="",sub="",main="Marker clustering on TOM based dissimilarity", labels=FALSE, hang=0.04)

# module identification using dynamic tree cut
minModuleSize=1
dynamicMods=cutreeDynamic(dendro=geneTree, distM=dissTOM, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize)
table(dynamicMods)

# convert numeric labels to colors
dynamicColors=labels2colors(dynamicMods)
table(dynamicColors)

# plot dendrograms and colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree,dynamicColors,"Dynamic Tree Cut", dendroLabels = FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05, main="Gene dendrogram and module colors")

# merging modules whose expression profiles are similar
MEList = moduleEigengenes(MDDinfF,colors=dynamicColors)
MEs = MEList$eigengenes
MEDiss = 1-cor(MEs)
METree = hclust(as.dist(MEDiss), method="average")
sizeGrWindow(7,6)
plot(METree, main="Clustering of module eigengenes", xlab="", sub="")

# choose height cut
MEDISSThres=0.25
merge=mergeCloseModules(MDDinfF,dynamicColors,cutHeight=MEDISSThres,verbose=3)
mergedColors=merge$colors
mergedMEs = merge$newMEs

sizeGrWindow(12,9)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),c("Dynamic Tree Cut", "Merged dynamic"), dendroLabels=FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05)

# save
moduleColors=mergedColors
colorOrder=c("grey",standardColors(50))
moduleLabels=match(moduleColors,colorOrder)-1
MEs=merge$newMEs
save(MEs, moduleLabels,moduleColors,geneTree,file="02-MDDinfF-networkConstruction.RData")
```

## Module output
```{r}
MDDFmodules = data.frame(markers = names(MDDinfF),
                        module = moduleColors)

write.csv(MDDFmodules, file = "MDDF.csv")
```

# Control male
## 0 setting up r session

```{r}
# load libraries
library(WGCNA)
options(stringsAsFactors = FALSE)
enableWGCNAThreads()

# load data saved in first step
lnames = load(file = "/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Sex/sex.RData")
lnames 
```


## 2a choosing the soft thresholding power: analysis of network topology

```{r}
# choose a set of soft thresholding powers
powers=c(c(1:10), seq(from=12, to=40, by=2))

# call network topology analysis function
sft = pickSoftThreshold(CntrlinfM, powerVector = powers, verbose = 5)

# plot the result
sizeGrWindow(9,5)
par(mfrow=c(1,2))
cex1=0.9

# scale free topology fit index as a function of the soft thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```

## 2b one step network construction and module detection

```{r}
# automatic construction is not working -- see below for step by step construction
net = blockwiseModules(CntrlinfM, power = 8,
                       TOMType = "unsigned", minModuleSize = 1,
                       reassignThreshold = 0, mergeCutHeight = 0.25,
                       numericLabels = TRUE, pamRespectsDendro = FALSE,
                       saveTOMs = TRUE,
                       saveTOMFileBase = "TOM", 
                       verbose = 3)

# module sizes and number of modules
table(net$colors)
```

create a table of soft thresholding powers

```{r}
# Initialize an empty list to store the lengths of net$MEs for each power
MEs_lengths <- numeric(20)

# Loop through powers from 1 to 20
for (power in 1:20) {
  # Run blockwiseModules for the current power
  net <- blockwiseModules(MDDinfF, power = power,
                           TOMType = "unsigned", minModuleSize = 1,
                           reassignThreshold = 0, mergeCutHeight = 0.25,
                           numericLabels = TRUE, pamRespectsDendro = FALSE,
                           saveTOMs = TRUE,
                           saveTOMFileBase = paste0("TOM_", power), 
                           verbose = 3)
  
  # Store the length of net$MEs for the current power
  MEs_lengths[power] <- length(net$MEs)
}

# Create a data frame to store the lengths of net$MEs for each power
MEs_lengths_table <- data.frame(Power = 1:20, Length_of_MEs = MEs_lengths)

# Display the table
print(MEs_lengths_table)
```

We plot the hierarchical clustering dendrogram for module identification.


```{r}
# open a graphics window
sizeGrWindow(12, 9)
# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                    "Module colors",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
```

## 2c step by step construction
```{r}
# calculate adjacencies
softPower=8
adjacency=adjacency(CntrlinfM,power=softPower)

# topological overlap matrix
TOM=TOMsimilarity(adjacency)
dissTOM=1-TOM

# clustering using TOM
geneTree=hclust(as.dist(dissTOM),method="average")
sizeGrWindow(12,9)
plot(geneTree,xlab="",sub="",main="Marker clustering on TOM based dissimilarity", labels=FALSE, hang=0.04)

# module identification using dynamic tree cut
minModuleSize=1
dynamicMods=cutreeDynamic(dendro=geneTree, distM=dissTOM, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize)
table(dynamicMods)

# convert numeric labels to colors
dynamicColors=labels2colors(dynamicMods)
table(dynamicColors)

# plot dendrograms and colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree,dynamicColors,"Dynamic Tree Cut", dendroLabels = FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05, main="Gene dendrogram and module colors")

# merging modules whose expression profiles are similar
MEList = moduleEigengenes(CntrlinfM,colors=dynamicColors)
MEs = MEList$eigengenes
MEDiss = 1-cor(MEs)
METree = hclust(as.dist(MEDiss), method="average")
sizeGrWindow(7,6)
plot(METree, main="Clustering of module eigengenes", xlab="", sub="")

# choose height cut
MEDISSThres=0.25
merge=mergeCloseModules(CntrlinfM,dynamicColors,cutHeight=MEDISSThres,verbose=3)
mergedColors=merge$colors
mergedMEs = merge$newMEs

sizeGrWindow(12,9)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),c("Dynamic Tree Cut", "Merged dynamic"), dendroLabels=FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05)

# save
moduleColors=mergedColors
colorOrder=c("grey",standardColors(50))
moduleLabels=match(moduleColors,colorOrder)-1
MEs=merge$newMEs
save(MEs, moduleLabels,moduleColors,geneTree,file="02-CntrlInfM-networkConstruction.RData")
```

## Module output
```{r}
CntrlMmodules = data.frame(markers = names(CntrlinfM),
                        module = moduleColors)

write.csv(CntrlMmodules, file = "CntrlM.csv")
```

# Control female 
## 0 setting up r session

```{r}
# load libraries
library(WGCNA)
options(stringsAsFactors = FALSE)
enableWGCNAThreads()

# load data saved in first step
lnames = load(file = "/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Sex/sex.RData")
lnames 
```


## 2a choosing the soft thresholding power: analysis of network topology

```{r}
# choose a set of soft thresholding powers
powers=c(c(1:10), seq(from=12, to=40, by=2))

# call network topology analysis function
sft = pickSoftThreshold(CntrlinfF, powerVector = powers, verbose = 5)

# plot the result
sizeGrWindow(9,5)
par(mfrow=c(1,2))
cex1=0.9

# scale free topology fit index as a function of the soft thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```

## 2b one step network construction and module detection

```{r}
# automatic construction is not working -- see below for step by step construction
net = blockwiseModules(CntrlinfF, power = 8,
                       TOMType = "unsigned", minModuleSize = 1,
                       reassignThreshold = 0, mergeCutHeight = 0.25,
                       numericLabels = TRUE, pamRespectsDendro = FALSE,
                       saveTOMs = TRUE,
                       saveTOMFileBase = "TOM", 
                       verbose = 3)

# module sizes and number of modules
table(net$colors)
```

create a table of soft thresholding powers

```{r}
# Initialize an empty list to store the lengths of net$MEs for each power
MEs_lengths <- numeric(20)

# Loop through powers from 1 to 20
for (power in 1:20) {
  # Run blockwiseModules for the current power
  net <- blockwiseModules(MDDinfF, power = power,
                           TOMType = "unsigned", minModuleSize = 1,
                           reassignThreshold = 0, mergeCutHeight = 0.25,
                           numericLabels = TRUE, pamRespectsDendro = FALSE,
                           saveTOMs = TRUE,
                           saveTOMFileBase = paste0("TOM_", power), 
                           verbose = 3)
  
  # Store the length of net$MEs for the current power
  MEs_lengths[power] <- length(net$MEs)
}

# Create a data frame to store the lengths of net$MEs for each power
MEs_lengths_table <- data.frame(Power = 1:20, Length_of_MEs = MEs_lengths)

# Display the table
print(MEs_lengths_table)
```

We plot the hierarchical clustering dendrogram for module identification.


```{r}
# open a graphics window
sizeGrWindow(12, 9)
# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                    "Module colors",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
```

## 2c step by step construction
```{r}
# calculate adjacencies
softPower=8
adjacency=adjacency(CntrlinfF,power=softPower)

# topological overlap matrix
TOM=TOMsimilarity(adjacency)
dissTOM=1-TOM

# clustering using TOM
geneTree=hclust(as.dist(dissTOM),method="average")
sizeGrWindow(12,9)
plot(geneTree,xlab="",sub="",main="Marker clustering on TOM based dissimilarity", labels=FALSE, hang=0.04)

# module identification using dynamic tree cut
minModuleSize=1
dynamicMods=cutreeDynamic(dendro=geneTree, distM=dissTOM, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize)
table(dynamicMods)

# convert numeric labels to colors
dynamicColors=labels2colors(dynamicMods)
table(dynamicColors)

# plot dendrograms and colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree,dynamicColors,"Dynamic Tree Cut", dendroLabels = FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05, main="Gene dendrogram and module colors")

# merging modules whose expression profiles are similar
MEList = moduleEigengenes(CntrlinfF,colors=dynamicColors)
MEs = MEList$eigengenes
MEDiss = 1-cor(MEs)
METree = hclust(as.dist(MEDiss), method="average")
sizeGrWindow(7,6)
plot(METree, main="Clustering of module eigengenes", xlab="", sub="")

# choose height cut
MEDISSThres=0.25
merge=mergeCloseModules(CntrlinfF,dynamicColors,cutHeight=MEDISSThres,verbose=3)
mergedColors=merge$colors
mergedMEs = merge$newMEs

sizeGrWindow(12,9)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),c("Dynamic Tree Cut", "Merged dynamic"), dendroLabels=FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05)

# save
moduleColors=mergedColors
colorOrder=c("grey",standardColors(50))
moduleLabels=match(moduleColors,colorOrder)-1
MEs=merge$newMEs
save(MEs, moduleLabels,moduleColors,geneTree,file="02-CntrlInfF-networkConstruction.RData")
```

## Module output
```{r}
CntrlFmodules = data.frame(markers = names(CntrlinfF),
                        module = moduleColors)

write.csv(CntrlFmodules, file = "CntrlF.csv")
```

