---
title: "3 Depression"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MDD
## 0 set up r session

```{r}
library(WGCNA)
options(stringsAsFactors=FALSE)
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Depression/MDDinfDep.RData")
lnames
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/2 module detection/02-MDD-networkConstruction.RData")
lnames
```

## 3a quantifying module trait associations

```{r}
# define numbers of genes and samples
nMarkers=ncol(MDDinfData5)
nSamples=nrow(MDDinfData5)

# recalculate MEs with color labels
MEs0=moduleEigengenes(MDDinfData5,moduleColors)$eigengenes
MEs=orderMEs(MEs0)
moduleTraitCor=cor(MEs,MDDdep2,use="p")
moduleTraitPvalue=corPvalueStudent(moduleTraitCor,nSamples)

# display correlations and p values
textMatrix = paste(signif(moduleTraitCor,2), "\n(",
                   signif(moduleTraitPvalue,1),")",sep="")
dim(textMatrix)=dim(moduleTraitCor)
par(mar=c(6,8.5,3,3))

labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = c("CGI", "GAD7", "MADRS","QIDS"),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```

## 3b gene relationship to trait and important modules: Gene Significance and Module Membership

Quantify associations of individual inflammatory markers with trait of interest by defining Gene Significance as the absolute value of the correlation between inflammatory marker and the trait. For each module we also define a quantitative measure of module membership MM as the correlation of the module eigengene and the gene expression profile. This allows us to quantify the similarity of all genes on the array to every module.

```{r}
ef = as.data.frame(MDDdep2$MADRS_TOT_PRO_RATED);
names(ef) = "MADRS"

# names (colors) of the modules
modNames = substring(names(MEs), 3)

geneModuleMembership = as.data.frame(cor(MDDdep2, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneTraitSignificance = as.data.frame(cor(MDDdep2, ef, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(ef), sep="");
names(GSPvalue) = paste("p.GS.", names(ef), sep="");

```

## 3c intramodular analysis: identifying genes with high GS and MM

See if GS and MM are highly correlated, which means that markers highly significantly associated with a trait are the most important elements of modules associated with the trait.

```{r}
module = "blue"
column = match(module, modNames);
moduleGenes = moduleColors==module;

sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Marker significance for executive",
                   main = paste("Module membership vs. marker significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

```
## visualize

visualize network of eigengenes

```{r}
# Recalculate module eigengenes
MEs = moduleEigengenes(MDDinfData5, moduleColors)$eigengenes
# Isolate weight from the clinical traits
ef = as.data.frame(MDDdep2$MADRS_TOT_PRO_RATED);
names(ef) = "MADRS"
# Add the weight to existing module eigengenes
MET = orderMEs(cbind(MEs, ef))
# Plot the relationships among the eigengenes and the trait
sizeGrWindow(5,7.5);
par(cex = 0.9)
plotEigengeneNetworks(MET, "", marDendro = c(0,4,1,2), marHeatmap = c(3,4,1,2), cex.lab = 0.8, xLabelsAngle
= 90)
```

# Control
## 0 set up r session

```{r}
library(WGCNA)
options(stringsAsFactors=FALSE)
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/1 data input/Depression/CntrlinfDep.RData")
lnames
lnames=load(file="/Users/lucyhui/Downloads/SUDS/Individual/2 module detection/02-Control-networkConstruction.RData")
lnames
```

## 3a quantifying module trait associations

```{r}
# define numbers of genes and samples
nMarkers=ncol(CntrlinfData5)
nSamples=nrow(CntrlinfData5)

# recalculate MEs with color labels
MEs0=moduleEigengenes(CntrlinfData5,moduleColors)$eigengenes
MEs=orderMEs(MEs0)
moduleTraitCor=cor(MEs,Cntrldep2,use="p")
moduleTraitPvalue=corPvalueStudent(moduleTraitCor,nSamples)

# display correlations and p values
textMatrix = paste(signif(moduleTraitCor,2), "\n(",
                   signif(moduleTraitPvalue,1),")",sep="")
dim(textMatrix)=dim(moduleTraitCor)
par(mar=c(6,8.5,3,3))

labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = c("CGI", "GAD7", "MADRS","QIDS"),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```

## 3b gene relationship to trait and important modules: Gene Significance and Module Membership

Quantify associations of individual inflammatory markers with trait of interest by defining Gene Significance as the absolute value of the correlation between inflammatory marker and the trait. For each module we also define a quantitative measure of module membership MM as the correlation of the module eigengene and the gene expression profile. This allows us to quantify the similarity of all genes on the array to every module.

```{r}
ef = as.data.frame(Cntrldep2$MADRS_TOT_PRO_RATED);
names(ef) = "MADRS"

# names (colors) of the modules
modNames = substring(names(MEs), 3)

geneModuleMembership = as.data.frame(cor(Cntrldep2, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneTraitSignificance = as.data.frame(cor(Cntrldep2, ef, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(ef), sep="");
names(GSPvalue) = paste("p.GS.", names(ef), sep="");

```

## 3c intramodular analysis: identifying genes with high GS and MM

See if GS and MM are highly correlated, which means that markers highly significantly associated with a trait are the most important elements of modules associated with the trait.

```{r}
module = "grey"
column = match(module, modNames);
moduleGenes = moduleColors==module;

sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Marker significance for executive",
                   main = paste("Module membership vs. marker significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

```
