---
title: "1 CNSVS"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2. CNSVS

## 2.a MDD

The baseline MDD CNSVS data is contained in the file `CBN01_CNSVS_DATA_Z3_01_V01_TRTMT.csv` that comes with CAN-BIND-1.

```{r}
# load package
library(WGCNA)
# do not omit
options(stringsAsFactors = FALSE)
# read data set
MDDcnsvs = read.csv("/Users/lucyhui/Downloads/SUDS/Individual/1 data input/CNSVS/CBN01_CNSVS_DATA_Z3_01_V01_TRTMT.csv")
```

`MDDcnsvs` contains 211 subjects with 99 variables. Each row corresponds to a patient and column to a measurement. We now remove auxiliary data and only include clincal domain of the CNS vital sign test.

```{r}
MDDcnsvs1 = as.data.frame(MDDcnsvs[,c(1, 14:28)])
```

Form a data frame analogous to expression data that will hold the clinical traits. Recall that N = 195.

```{r}
traitRows = match(rownames(MDDinfData5), MDDcnsvs1$SUBJLABEL)
MDDcnsvs2 = MDDcnsvs1[traitRows, -1]
rownames(MDDcnsvs2) = MDDcnsvs1[traitRows, 1]
```

We visualize how the clinical traits relate to the sample dendrogram.

```{r}
# white means low value, red means high value, grey means missing entry
sampleTree2 = hclust(dist(MDDinfData5), method = "average")
traitColors = numbers2colors(MDDcnsvs2, signed = FALSE)
plotDendroAndColors(sampleTree2, traitColors, groupLabels = names(MDDcnsvs2), main = "Sample dendrogram and trait heatmap")
```

Save results for the next step.

```{r}
save(MDDinfData5, MDDcnsvs2, file = "MDDinfCNSVS1.RData")
```

## 2.a Control

The baseline MDD CNSVS data is contained in the file `CBN01_CNSVS_DATA_Z3_01_V01_CNTRL.csv` that comes with CAN-BIND-1.

```{r}
# load package
library(WGCNA)
# do not omit
options(stringsAsFactors = FALSE)
# read data set
Cntrlcnsvs = read.csv("/Users/lucyhui/Downloads/SUDS/Individual/1 data input, cleaning, pre-processing/CNSVS/CBN01_CNSVS_DATA_Z3_01_V01_CNTRL.csv")
```

```{r}
Cntrlcnsvs1 = as.data.frame(Cntrlcnsvs[,c(1, 14:28)])
```

Form a data frame analogous to expression data that will hold the clinical traits.

```{r}
traitRows = match(rownames(CntrlinfData5), Cntrlcnsvs1$SUBJLABEL)
Cntrlcnsvs2 = Cntrlcnsvs1[traitRows, -1]
rownames(Cntrlcnsvs2) = Cntrlcnsvs1[traitRows, 1]
```

We visualize how the clinical traits relate to the sample dendrogram.

```{r}
# white means low value, red means high value, grey means missing entry
sampleTree2 = hclust(dist(CntrlinfData5), method = "average")
traitColors = numbers2colors(Cntrlcnsvs2, signed = FALSE)
plotDendroAndColors(sampleTree2, traitColors, groupLabels = names(Cntrlcnsvs2), main = "Sample dendrogram and trait heatmap")
```

Save results for the next step.

```{r}
save(CntrlinfData5, Cntrlcnsvs2, file = "CNTRLinfCNSVS1.RData")
```

